Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 4.dev

### 2025-11-15 review/review-response-01.tex

First, see the answer above to review \#1 as a partial response to your
question. The two fall modes are not technically the same and do not simply
differ in degree, but were necessary given the nature of the experimental
design. We did not record the category of the fall mode, so it is not possible
for us to investigate whether having a tenary factor instead of a binary one
would reveal any more insight. We added text in Section 2.4 to make this clear.

\begin{displayquote}
  Section 2.6: Statistics on page 11

  The authors state that "cluster-mean centering showed there to be no
  variation between participants," which allowed authors to use a simple
  single-level logistic regression model rather than a multilevel model. This
  assertion is a foundational premise for the authors' entire analysis.
  Therefore, the authors are required to present the statistical testing
  process for cluster mean centering.
\end{displayquote}

In Section 2.6, we have included the maximum intraclass correlation (ICC) of
the models that include the participant as a random effect, which is 3\%.
Including this random effect does lower the p-values by a small amount, since a
bit more variation is accounted for, but we decided to use the more
consertative approach and not include this random effect. The participants do
not have appreciable variation in how they respond to the perturbations after
the data is normalized per participant. This lack of variation was also
observed in the independent dataset collected in Marten Haitjema's MSc thesis.

### 2025-11-15 main.tex
Simplified units.

\subsection{Balance Assist Control}
%
We use a forward speed \(v\) gain scheduled proportional roll angular rate
feedback controller to stabilize the bicycle. In the speed range tested, the
commanded steer torque \(T^\textrm{m}_\delta\) from the steer motor follows the
control law
%
\begin{align}
  T^\textrm{m}_\delta = -k_{\dot{\phi}}\dot{\phi} = \kappa(v_{stable} - v)\dot{\phi}
  \label{eq:implemented-controller}
\end{align}
%
where \(v_{\textrm{stable}} = 4.7~\si{\meter\per\second}\) is approximately the
average stable speed predicted from the open-loop bicycle rigid-rider dynamics.
We use the gain factor \(\kappa=3.9\) (low) and \(\kappa=5.2\) (high) to tune
the gain magnitude during the experiments. Figure~\ref{fig:eigenvalues} shows
the stabilization effect of the controller via the eigenmodes of the linear
system. Geometric and inertial parameters for these plots were estimated with
the methods presented in \citet{Moore2012} and software packages
BicycleParameters~1.1.1~\citep{Moore2024} and Yeadon~1.5.0~\citep{Dembia2015}
and are shown in Appendix~\ref{sec:physical-parameters}.

@@

\subsection{Perturbations}
%
We apply longitudinal forces to the ends of each handlebar from earth-fixed
posts using an adapted Bump'Em system~\citep{Tan2020} which we arrange with
four EC-90 Flat motors (Maxon Group, Switzerland) working in cooperation,
Figure~\ref{fig:setup-diagram}. The four motors are programmed to apply a light
force at all times to keep the ropes taught and to track a commanded force
profile using a PID controller running on an Arduino Mega 2560 microprocessor
(Arduino, Italy). We measure the force applied by each motor at the handlebar
via four inline S-style load cells (Bosche GmbH, Damme, Germany). All were
rated for 250~\si{\newton} except the left rear load cell which was rated for
500~\si{\newton}. The commanded force profiles are designed to apply an
external pulsive torque to the front frame at magnitudes varying from
\SIrange{16}{160}{\newton\meter}. The four motors are arranged at the four
corners of a 1~\si{\meter} wide by 2~\si{\meter} long treadmill (Bonte
Technology B.V., Zwolle, The Netherlands) that can reach speeds up to
18~\si{\kilo\meter\per\hour}. The general design of this perturbation system is
described in detail in Van De Velde's MSc thesis~\citep{vandeVelde2022a} and
the physical arrangement is shown in Figure~\ref{fig:participant-in-set-up}.
Our modifications relative to Van De Velde's design include simplifying the
Bump'Em motor controller with an inexpensive microcontroller and the use of a
simpler non-active safety harness.
%
% NOTE : I get a compile error without the \protect calls here, not sure why.
\begin{figure}
  \centering
  \subcaptionbox{Top view diagram of bicycle handlebar perturbation
    system.\protect\label{fig:setup-diagram}}
    {\includegraphics[height=65mm]{figures/setup-diagram.png}}
  \subcaptionbox{A participant on the bicycle in the safety harness with the
    Bump'Em motor lines attached to the ends of the
    handlebars.\protect\label{fig:participant-in-set-up}}
    {\includegraphics[height=65mm]{figures/participant-in-set-up.png}}
  \caption{Diagrams showing the Bump'Em system arranged to apply handlebar
    perturbations and the various system components.}
  \label{fig:experimental-setup}
\end{figure}

>>>>

\subsection{Balance Assist Control}
%
We use a forward speed \(v\) gain scheduled proportional roll angular rate
feedback controller to stabilize the bicycle. In the speed range tested, the
commanded steer torque \(T^\textrm{m}_\delta\) from the steer motor follows the
control law
%
\begin{align}
  T^\textrm{m}_\delta = -k_{\dot{\phi}}\dot{\phi} = \kappa(v_{stable} - v)\dot{\phi}
  \label{eq:implemented-controller}
\end{align}
%
where \(v_{\textrm{stable}} = 4.7~\si{\mps}\) is approximately the average
stable speed predicted from the open-loop bicycle rigid-rider dynamics.  We use
the gain factor \(\kappa=3.9\) (low) and \(\kappa=5.2\) (high) to tune the gain
magnitude during the experiments. Figure~\ref{fig:eigenvalues} shows the
stabilization effect of the controller via the eigenmodes of the linear system.
Geometric and inertial parameters for these plots were estimated with the
methods presented in \citet{Moore2012} and software packages
BicycleParameters~1.1.1~\citep{Moore2024} and Yeadon~1.5.0~\citep{Dembia2015}
and are shown in Appendix~\ref{sec:physical-parameters}.

@@

\subsection{Perturbations}
%
We apply longitudinal forces to the ends of each handlebar from earth-fixed
posts using an adapted Bump'Em system~\citep{Tan2020} which we arrange with
four EC-90 Flat motors (Maxon Group, Switzerland) working in cooperation,
Figure~\ref{fig:setup-diagram}. The four motors are programmed to apply a light
force at all times to keep the ropes taught and to track a commanded force
profile using a PID controller running on an Arduino Mega 2560 microprocessor
(Arduino, Italy). We measure the force applied by each motor at the handlebar
via four inline S-style load cells (Bosche GmbH, Damme, Germany). All were
rated for 250~\si{\newton} except the left rear load cell which was rated for
500~\si{\newton}. The commanded force profiles are designed to apply an
external pulsive torque to the front frame at magnitudes varying from
\SIrange{16}{160}{\newton\meter}. The four motors are arranged at the four
corners of a 1~\si{\meter} wide by 2~\si{\meter} long treadmill (Bonte
Technology B.V., Zwolle, The Netherlands) that can reach speeds up to
18~\si{\kph}. The general design of this perturbation system is described in
detail in Van De Velde's MSc thesis~\citep{vandeVelde2022a} and the physical
arrangement is shown in Figure~\ref{fig:participant-in-set-up}.  Our
modifications relative to Van De Velde's design include simplifying the Bump'Em
motor controller with an inexpensive microcontroller and the use of a simpler
non-active safety harness.
%
% NOTE : I get a compile error without the \protect calls here, not sure why.
\begin{figure}
  \centering
  \subcaptionbox{Top view diagram of bicycle handlebar perturbation
    system.\protect\label{fig:setup-diagram}}
    {\includegraphics[height=65mm]{figures/setup-diagram.png}}
  \subcaptionbox{A participant on the bicycle in the safety harness with the
    Bump'Em motor lines attached to the ends of the
    handlebars.\protect\label{fig:participant-in-set-up}}
    {\includegraphics[height=65mm]{figures/participant-in-set-up.png}}
  \caption{Diagrams showing the Bump'Em system arranged to apply handlebar
    perturbations and the various system components.}
  \label{fig:experimental-setup}
\end{figure}
